Untwisting the Tetrahelix (v0.5)

### 2017-02-17 doc/TetrahelixUntwist.tex
improvements in figures and others

\begin{figure}[H] %float with two figures
  \centering
     \includegraphics[width=0.4\textwidth]{figures/Tetrahelix0.png}
     \caption{Regular Tetrahelix}
\end{figure}

BC helix does not rest on a plane in a simple way. It is convenient to be able to ``untwist'' it and form a tetrahelix space frame that
has a flat planar surface. By making length changes in a certain way, we can untwist a tetrahelix to form a \emph{tetrabeam} which
has planar faces and has, for example, an equilateral triangular profile.

@@

From Robert Gray's site, repeating formula by H.S.M. Coxeter:
\[
V(n) =
\left [
  \begin{tabular}{c}
   $ r_{bc} \cos(n \theta) $\\
   $ r_{bc} \sin(n \theta) $\\
   $ n h_{bc}  $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} $\\
 $ h_{bc} = 1/\sqrt{10} $ \\
 $ \theta = \arccos(-2/3) $ \\
  \end{tabular}      
\]
where $n$ represents each integer numbered node in succession.

@@

Such a helix can be written:
\[
H_{BCcolored}(n,c) =
\left [
  \begin{tabular}{c}
   $ r_{bc}  \cos((3 \theta - 2 \pi)n + c  \theta $\\
   $ r_{bc} \sin((3 \theta - 2 \pi)n + c  \theta $\\
   $ (n + c/3) 3  h_{bc} $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} $\\
 $ h_{bc} = 1/\sqrt{10} $ \\
 $ \theta = \arccos(-2/3) $ \\
  \end{tabular}      
  \]

@@

 \begin{figure}[H]
     \centering
     \includegraphics[width=0.9\textwidth]{figures/TetrahelixColoringDiagram.png}
     \caption{Coloring of an (untwisted) Tetrabeam}
 \end{figure}

@@

\[
H_{BCsymmmetric}(n,c) =
\left [
  \begin{tabular}{c}
   $ r  \cos(\rho_{bc} \kappa  + c 2 \pi /3) $\\
   $ r  \sin(\rho_{bc} \kappa  + c 2 \pi /3) $\\
   $ \kappa 3  h_{bc} $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $\kappa = n + c/3$ \\
    $\rho_{bc} = (3 \theta - 2 \pi)$ \\
    $ \theta = \arccos(-2/3) $
    $ h_{bc} = 1/\sqrt{10} $ \\    
  \end{tabular}      
\]

@@

\begin{align*}
  1^2 &= d^2 + (2 r \sin{ \rho / 2})^2 \\
  1 &= d^2 + 4 r^2 (\sin{ \rho / 2})^2 \\
  d^2 &= 1 - 4 r^2 (\sin{ \rho / 2})^2   
\end{align*}

@@

Rewriting our formulation in terms of $\rho$:
\[
H_{general}(n,c,\rho,d_{\rho},r_{\rho}) =
\left [
  \begin{tabular}{c}
   $ r_{\rho} \cos(\rho \kappa + c 2 \pi /3) $\\
   $ r_{\rho}  \sin(\rho \kappa + c 2 \pi /3) $\\
   $ d_{\rho} \kappa $
  \end{tabular}
  \right ],
\text{where:}
\begin{tabular}{c}
  $   1 = d_{\rho}^2 + 4 r_{\rho}^2 (\sin{ \rho / 2})^2 $ \\
    $\kappa = n + c/3$ \\
  \end{tabular}      
\]

@@

Therefore we decrease the minimax length
of the whole system as we increase the radius
up unto the point that the shorter, one-hop distance is equal to the rail-length ($1$).
Therefore, to optimize the whole system so long as $\rho < \rho_{bc}$,
we equate one-hop to $1$ to find the optimum radius:

@@

Note: $ d_{opt} $ is computed by $\rho, r_{opt}$ via the rail equation \eqref{eqrhoopt}.

Therefore, by computing $r_{opt}$ as a function of $\rho$ from this equation, we can construct minimax optimal tetrahelix for an $\rho < \rho_{bc}$.

@@

However, choosing $d = 1$ and $\rho = 0$ does not fix the radius.
We would like to choose the radius so that the
minimax difference between the lengths is as small as possible.

 \begin{figure}[H]
     \centering
     \includegraphics[width=0.9\textwidth]{figures/EquilateralDiagram.png}
     \caption{Coloring of an (untwisted) Tetrabeam}
 \end{figure}

@@

When $\rho = 0$, our formulae for distances are particular simple. Let $a$ be the length of the equilateral triangle
forming the beam cross section. Then by geometry $r = a \sqrt{3}$.
\begin{align*}
  \text{rail} &=  1 \\
  \text{one-hop} &= \sqrt{\frac{1}{9} + 3r^2}\\
\text{two-hops}    &= \sqrt{\frac{4}{9} + 3r^2}
\end{align*}
Computing the derivative of two-hops - one-hop:
\begin{align*}
 \frac{\partial \text{two-hops} - \text{one-hop}}{r} &= \frac{\partial \sqrt{4/9 + 3r^2} - \sqrt{1/9 + 3r^2}}{\partial r} \\
  \frac{\partial \text{two-hops} - \text{one-hop}}{r} &= 9r (\frac{1}{\sqrt{27r^2 + 4}} - \frac{1}{\sqrt{ 27r^2 + 1}}) 
\end{align*}
which is negative for all $r> 0$. Thererfore two-hop - one-hop
is minimized when we maximize $r$, up until one-hop = rail.

@@

@@

We observe that by thinking of the straight lines of the
Equitetrabeam as a degenerate helix of zero curvature,
it should be possible to define a smoothly varying continuum between the Boerdijk--Coxeter helix and the Equitetrabeam and every
curvature and torsion between the two.

We seek to unify this with degenerate helix formula for the equitetrabeam:
\[
H_{etb}(n,c) =
\left [
  \begin{tabular}{c}
   $ r_{etb}  \cos( 0 \cdot \kappa  + c 2 \pi /3) $\\
   $ r_{etb}  \sin( 0 \cdot \kappa  + c 2 \pi /3) $\\
   $ \kappa d_{etb} $
  \end{tabular}
\right ],
\text{where:}
  \begin{tabular}{c}
 $\kappa = n + c/3$ \\
    $\rho_{bc} = (3 \theta - 2 \pi)$ \\
   $ \theta = \arccos(-2/3) $
  \end{tabular}      
\]
where $ r_{etb} = \sqrt{\frac{8}{27}}$, $d_{etb} = 1$,

\[
H_{continuum}(n,c,\lambda) =
\left [
  \begin{tabular}{c}
   $ r_{\lambda} \cos(\phi_{\lambda} \kappa + c 2 \pi /3) $\\
   $ r_{\lambda}  \sin(\phi_{\lambda} \kappa + c 2 \pi /3) $\\
   $ d_{\lambda} \kappa $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
    $\kappa = n + c/3$ \\
    $ r_{opt} =  \text{is used : \eqref{eqrhoopt}} $ \\    
    $ \phi_{\lambda} =  \lambda \rho_{bc}  + 0 $\\
    $\rho_{bc} = (3 \theta - 2 \pi)$ \\
   $ \theta = \arccos(-2/3) $
  \end{tabular}      
\]
A value of $\lambda = 0$ generates the equitetrabeam, and $\lambda = 1$ generates the Boerdijk--Coxeter helix, and every
value $\lambda \in [-1,1]$ generates an attractive structure which if physically realized is an inherently rigid structure
with member lengths of no greater disparity than $16\%$.
In the sense of structural engineering, it would be a relatively strong space frame. When $\lvert \lambda \rvert > 1$, tetrahelices
are still produced, but as $\lvert \lambda \rvert >> 2$ these tetrahelices probably cease to be interesting to mechanical engineers,
because their ability to withstand axial forces seem to diminish significantly.

@@

Negative values of $\lambda$ generate clockwise tetrahelices; positive generate counter-clockwise helices.

Furthermore, this formula allows one to design the pitch (in edge length units) as a function of $\lambda$ of the helix (along one rail)
where $\lambda \neq 0$:
\[
p(\lambda) = 2 \pi  \cdot \frac{(\frac{8}{27} -1) \lambda +1)}{ \lambda  \rho_{bc} }
\]

We can then create a continuum of radius, length, rho to place into our general equation
to create continuum of helices between the BC and equitetrabeam. This continuum is 
minimax-length optimal at its endpoints, and may be optimal throughout. This untwisting is accomplished
only by changing the length of the two-hop member.

@@

@@

>>>>

\begin{figure}[H] %float with two figures
  \label{series}
  \centering
     \includegraphics[width=0.95\textwidth]{figures/TetrahelixSeries.png}
     \caption{A Series of Tetrahelices, BC Helix in Foreground.}
\end{figure}

BC helix does not rest on a plane in a simple way. It is convenient to be able to ``untwist'' it and form a tetrahelix space frame that
has a flat planar surface. By making length changes in a certain way, we can untwist a tetrahelix to form a \emph{tetrabeam} which
has planar faces and has, for example, an equilateral triangular profile.
In Figure \ref{series}, the closest helix is the BC helix.

@@

From Robert Gray's site, repeating formula by H.S.M. Coxeter:
\begin{equation}
V(n) =
\left [
  \begin{tabular}{c}
   $ r_{bc} \cos(n \theta) $\\
   $ r_{bc} \sin(n \theta) $\\
   $ n h_{bc}  $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} $\\
 $ h_{bc} = 1/\sqrt{10} $ \\
 $ \theta = \arccos(-2/3) $ \\
  \end{tabular}      
\end{equation}
where $n$ represents each integer numbered node in succession.

@@

Such a helix can be written:
\begin{equation}
H_{BCcolored}(n,c) =
\left [
  \begin{tabular}{c}
   $ r_{bc}  \cos((3 \theta - 2 \pi)n + c  \theta $\\
   $ r_{bc} \sin((3 \theta - 2 \pi)n + c  \theta $\\
   $ (n + c/3) 3  h_{bc} $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} $\\
 $ h_{bc} = 1/\sqrt{10} $ \\
 $ \theta = \arccos(-2/3) $ \\
  \end{tabular}      
\end{equation}

@@

@@

\begin{equation}
H_{BCsymmmetric}(n,c) =
\left [
  \begin{tabular}{c}
   $ r  \cos(\rho_{bc} \kappa  + c 2 \pi /3) $\\
   $ r  \sin(\rho_{bc} \kappa  + c 2 \pi /3) $\\
   $ \kappa 3  h_{bc} $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $\kappa = n + c/3$ \\
    $\rho_{bc} = (3 \theta - 2 \pi)$ \\
    $ \theta = \arccos(-2/3) $ \\
    $ h_{bc} = 1/\sqrt{10} $ \\    
  \end{tabular}      
\end{equation}

@@

\begin{align*}
  1^2 &= d^2 + (2 r \sin{ \rho / 2})^2 \\
  1 &= d^2 + 4 r^2 (\sin{ \rho / 2})^2 \\
  d^2 &= 1 - 4 r^2 (\sin{ \rho / 2})^2    \numberthis  \label{railangle} \\
\end{align*}

@@

Rewriting our formulation in terms of $\rho$:
\begin{equation}
H_{general}(n,c,\rho,d_{\rho},r_{\rho}) =
\left [
  \begin{tabular}{c}
   $ r_{\rho} \cos(\rho \kappa + c 2 \pi /3) $\\
   $ r_{\rho}  \sin(\rho \kappa + c 2 \pi /3) $\\
   $ d_{\rho} \kappa $
  \end{tabular}
  \right ],
\text{where:}
\begin{tabular}{c}
  $   1 = d_{\rho}^2 + 4 r_{\rho}^2 (\sin{ \rho / 2})^2 $ \\
    $\kappa = n + c/3$ \\
  \end{tabular}      
\end{equation}

@@

Therefore we decrease the minimax length
of the whole system as we increase the radius
up unto the point that the shorter, one-hop distance is equal to the rail-length ($1$).
Therefore, to optimize the whole system so long as $\rho \leq \rho_{bc}$,
we equate one-hop to $1$ to find the optimum radius:

@@

Note: $ d_{opt} $ is computed by $\rho, r_{opt}$ via the rail angle equation \eqref{railangle}.

\begin{align*}
  d_{opt}^2 &= 1 - 4 \frac{2}{\sqrt{\frac{9}{2} \cdot ( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + cos(\rho)+ 8 }}^2 (\sin{ \rho / 2})^2   \\
  d_{opt}^2 &= 1 - 4 \frac{4}{\frac{9}{2} \cdot ( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + cos(\rho)+ 8 } (\sin{ \rho / 2})^2   \\
  d_{opt}^2 &= 1 - \frac{32}{9 ( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + cos(\rho)+ 8 } (\sin{ \rho / 2})^2   \\
  d_{opt}^2 &= 1 - \frac{8 sin^2(\rho) }{ cos(r) + 9( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + 8 }   \\
    d_{opt} &= \sqrt{1 - \frac{8 sin^2(\rho) }{ cos(r) + 9( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + 8 }}    \numberthis  \label{dopt}  \\      
\end{align*}
Therefore, by computing $r_{opt}$ as a function of $\rho$ from this equation, we can construct minimax optimal tetrahelix for an $\rho < \rho_{bc}$.

@@

However, choosing $d = 1$ and $\rho = 0$ we use Equation \eqref{eqrhoopt} to find the radius of 
optimal minimax difference.

@@

\begin{align*}
  \text{rail} &=  1 \\
  \text{one-hop} &= \sqrt{\frac{1}{9} + 3r^2}\\
\text{two-hops}    &= \sqrt{\frac{4}{9} + 3r^2}
\end{align*}
Computing the derivative of two-hops - one-hop:
\begin{align*}
 \frac{\partial \text{two-hops} - \text{one-hop}}{r} &= \frac{\partial \sqrt{4/9 + 3r^2} - \sqrt{1/9 + 3r^2}}{\partial r} \\
  \frac{\partial \text{two-hops} - \text{one-hop}}{r} &= 9r (\frac{1}{\sqrt{27r^2 + 4}} - \frac{1}{\sqrt{ 27r^2 + 1}}) 
\end{align*}
which is negative for all $r> 0$. Thererfore two-hop - one-hop
is minimized when we maximize $r$, up until one-hop = rail.

@@

In Figure \ref{series}, the furthest tetrahelix is the optimal equitetrabeam.

@@

We observe that by thinking of the straight lines of the
Equitetrabeam as a degenerate helix of zero curvature,
it should be possible to define a smoothly varying continuum between the Boerdijk--Coxeter helix and the Equitetrabeam and every
curvature and torsion between the two. It seems most convenient to define this continuum in terms
of the rail angle $\rho$, where $0 \leq \rho \leq \rho_{bc}$.

Now basic components of the helix, which are the radius $r$, the rate of rotation, and the rate of
axial growth can all be linearly interpolated with a parameter $\lambda$ between their high values (for the BC helix)
and low values (for the equitetrabeam).

\begin{align*}
\phi_{\lambda} &=  \lambda \cdot \rho_{bc}   \\
r_{\lambda}  &=  r_{opt}(\phi) \\
d_{\lambda} &=   d_{opt}(\phi) \\
\end{align*}
to create a formula that generates a continuum of tetrahedral structures:

\begin{equation}
H_{continuum}(n,c,\lambda) =
\left [
  \begin{tabular}{c}
   $ r_{\lambda} \cos(\phi_{\lambda} \kappa + c 2 \pi /3) $\\
   $ r_{\lambda}  \sin(\phi_{\lambda} \kappa + c 2 \pi /3) $\\
   $ d_{\lambda} \kappa $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
    $\kappa = n + c/3$ \\
    $ \phi_{\lambda} =  \lambda \rho_{bc}  $\\
    $\rho_{bc} = (3 \theta - 2 \pi)$ \\
   $ \theta = \arccos(-2/3) $
  \end{tabular}      
\end{equation}

A value of $\lambda = 0$ generates the equitetrabeam, and $\lambda = 1$ generates the Boerdijk--Coxeter helix, and every
value $\lambda \in [-1,1]$ generates an attractive structure which if physically realized is an inherently rigid structure
with member lengths of no greater disparity than $16\%$.
In the sense of structural engineering, it would be a relatively strong space frame. When $\lvert \lambda \rvert > 1$, tetrahelices
are still produced, but as $\lvert \lambda \rvert >> 2$ these tetrahelices probably cease to be interesting to mechanical engineers,
because their ability to withstand axial forces seem to diminish significantly.

@@

 Negative values of $\lambda$ generate clockwise tetrahelices; positive generate counter-clockwise helices.
 (Note: In fact there is something wrong with my math, negative lambda values do produces opposit chirality
 but the lengths are not the same as the positive values.  I need to track this down.)

Furthermore, this formula allows one to design the pitch (in edge length units) as a function of $\lambda$ of the helix (along one rail)
where $\lambda \neq 0$:
\[
p(\lambda) = 2 \pi  \cdot d_{opt}
\]

There is thus an optimal continuum of tetrahelices untwisting between BC and equitetrabeam.
Perhaps surprisingly, this untwisting is accomplished only by changing the length of the two-hop member.

@@

\section{Conclusion}

@@

\[
H_{continuum}(n,c,\lambda) =
\left [
  \begin{tabular}{c}
   $ r_{\lambda} \cos(\phi_{\lambda} \kappa + c 2 \pi /3) $\\
   $ r_{\lambda}  \sin(\phi_{\lambda} \kappa + c 2 \pi /3) $\\
   $ d_{\lambda} \kappa $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
    $\kappa = n + c/3$ \\
    $ r_{opt} =  \text{is used : \eqref{eqrhoopt}} $ \\    
    $ \phi_{\lambda} =  \lambda \rho_{bc}  + 0 $\\
    $\rho_{bc} = (3 \theta - 2 \pi)$ \\
   $ \theta = \arccos(-2/3) $
  \end{tabular}      
\]
